---
title: "p8105_hw3_lz2951"
author: "lz2951"
date: "2023-10-07"
output: github_document
---

# Import packages

```{r}
library(p8105.datasets)
library(tidyverse)
```

# Problem 1

## Dataset description

First, import dataset.

```{r}
data("instacart")
```

The dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row is a product from an order and each column is a variable. The 15 variables are: `r colnames(instacart)`. Take the observation `r instacart[1,]` as an example, it shows a `r instacart[1,]$product_name` with product id `r instacart[1,]$product_id` bought at `r instacart[1,]$order_hour_of_day` o'clock of the `r instacart[1,]$order_dow`th day of the week by a user with user id `r instacart[1,]$user_id`. This product purchase belongs to the order with order id `r instacart[1,]$order_id` and order sequence order `r instacart[1,]$order_number`, in which the order of this product to be added to the cart is `r instacart[1,]$add_to_cart_order`. It has been `r instacart[1,]$days_since_prior_order` days since the last order. The yogurt has been ordered by the user in the past, and belongs to the aisle `r instacart[1,]$aisle` with aisle id `r instacart[1,]$aisle_id` and department `r instacart[1,]$department` with department id `r instacart[1,]$department_id`. The evaluation set this order belongs in is `r instacart[1,]$eval_set`.

## Some key questions about dataset

- How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart |>
  group_by(aisle) |>
  summarize(orders_from_aisle = n()) |>
  nrow()
```

There are 134 aisles.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(orders_from_aisle = n()) |>
  mutate(
    aisle_rank_by_order = min_rank(desc(orders_from_aisle))
  ) |>
  filter(aisle_rank_by_order < 2)
```

The aisle with the most items are fresh vegetables.

- Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(orders_from_aisle = n()) |>
  filter(orders_from_aisle > 10000) |>
  ggplot(aes(x = fct_reorder(aisle, desc(orders_from_aisle)), y = orders_from_aisle)) +
  geom_col() +
  coord_flip() +
  labs(x = "aisle", y = "orders belonging to certain aisle")
```

- Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  group_by(aisle, product_name) |>
  summarize(number_of_orders = n()) |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  mutate(
    product_rank_in_aisle = min_rank(desc(number_of_orders))
  ) |>
  filter(product_rank_in_aisle < 4) |>
  arrange(aisle, product_rank_in_aisle) |>
  knitr::kable()
```

- Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(order_dow, product_name) |>
  summarize(mean_hour_of_the_day = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour_of_the_day
  ) |>
  rename(
    mean_hour_of_Sunday = 2,
    mean_hour_of_Monday = 3,
    mean_hour_of_Tuesday = 4,
    mean_hour_of_Wednesday = 5,
    mean_hour_of_Thusday = 6,
    mean_hour_of_Friday = 7,
    mena_hour_of_Saturday = 8
  ) |>
  knitr::kable()
```

# Problem 2

## Data cleaning

First, import dataset.

```{r}
data("brfss_smart2010")
```

Check and clean the variable names:

```{r}
colnames(brfss_smart2010)
brfss_smart2010 = janitor::clean_names(brfss_smart2010)
```

Focus on the "Overall Health" topic.

```{r}
brfss_smart2010 |>
  group_by(topic) |>
  summarize(n_topic = n()) |>
  filter(topic == "Overall Health")

brfss_smart2010_overall_health =
  brfss_smart2010 |>
    filter(topic == "Overall Health")
```

Include only responses from "Excellent" to "Poor":

```{r}
brfss_smart2010_overall_health |>
  group_by(response) |>
  summarize(n_response = n())
```

Organize responses as a factor taking levels ordered from "Poor" to "Excellent":

```{r}
brfss_smart2010_overall_health =
  brfss_smart2010_overall_health |>
    mutate(response = factor(brfss_smart2010_overall_health$response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

## Some questions about dataset

- In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010_overall_health |>
  group_by(year, locationabbr, locationdesc) |>
  summarize(n_location = n()) |>
  group_by(year, locationabbr) |>
  summarize(n_location_in_state = n()) |>
  filter(n_location_in_state >= 7, year %in% c(2002, 2010)) |>
  arrange(year) |>
  knitr::kable()
```

According to the resulting table, in 2002, states observed at 7 or more locations are as following:
CT, FL, MA, NC, NJ, PA. In 2010, those states are CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA.

- Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help). 

```{r}
brfss_smart2010_overall_health |>
  filter(response == "Excellent") |>
  select(year, state = locationabbr, data_value) |>
  group_by(year, state) |>
  summarize(average_data_value = mean(data_value, na.rm = TRUE)) |>
  ggplot(aes(x = year, y = average_data_value, group = state, color = state)) +
  geom_line() +
  xlab("Year") +
  ylab("Average of the data_value across locations within a state") +
  ggtitle("Plot of Average data_value over time within each state")
```

According to the resulting plot, DC had the highest average data_values in most of years between 2002 and 2010, while WV had the lowest average data_values in most of years.

- Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_smart2010_overall_health |>
  filter(year %in% c(2006, 2010), locationabbr == "NY") |>
  ggplot(aes(x = response, y = data_value)) +
  geom_violin(aes(fill = response), alpha = 0.5) +
  facet_grid(~year) +
  ggtitle("Plot Showing Distribution of data_value for Response in NY for the years 2006 and 2010")
```

According to the resulting plots, the distributions of data_value for response in NY are similar in the years of 2006 and 2010.

# Problem 3

## Load, tidy, merge and orgnize the data sets

Import data and exclude participants less than 21 years of age and those with missing demographic data.

```{r}
demographic =
  read_csv(file = "./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  drop_na() |>
  filter(age >= 21)

accelerometer =
  read_csv(file = "./data/nhanes_accel.csv") |>
  janitor::clean_names()
```

Encode data with reasonable variable classes.

```{r}
demographic =
  mutate(
    demographic, 
    sex = factor(sex, levels = c(1, 2), labels = c("male", "female")), 
    education = factor(education, levels = c(1, 2, 3), 
                       labels = c("Less than high school", 
                                  "High school equivalent", 
                                  "More than high school"),
                       ordered = TRUE)
  )
```

Merge two data sets.

```{r}
nhanes_data = 
  inner_join(demographic, accelerometer, by = "seqn")
```

## Analysis of total activity

```{r}
nhanes_data |>
  mutate(total_activity = rowSums(across(min1 : min1440))) |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth() +
  facet_grid(~education) +
  ggtitle("24-hour Total Activity over Age for Different Sex and Education Levels")
```

Based on the results plot, for participants with an education equivalent to or more than high school, female's total activity is higher than male's at almost all ages. For participants with an education less than high school, female's total activity is only higher than male's before the age of 40. In all educational levels, the total activity curves for both genders look similar in shape.

## Analysis of activity over the course of the day

```{r}
nhanes_data |>
  pivot_longer(
    min1 : min1440, names_to = "time_min", 
    values_to = "activity_min",
    names_prefix = "min",
    names_transform = list(time_min = as.integer)
  ) |>
  ggplot(aes(x = time_min, y = activity_min, color = sex)) +
  geom_point(shape = ".", alpha = 0.25) +
  facet_grid(~education) +
  geom_smooth() +
  xlab("time/min") +
  ylab("activity") +
  ggtitle("Plot of Inspection Activity over the Course of the Day")
```

According to the results, it can be observed that, regardless of the education level, the curves of activity over time course for both male and female are highly similar. This suggests that people's daily activity patterns tend to be consistent regardless of their education level or gender. In the Less than high school and High school equivalent education levels, the curves for both genders almost overlap. In the More than high school education level, female shows slightly higher activity level in non-sleep period of time.
